---
title: "Disparities in access to diagnosis and care in Blantyre, Malawi identified through enhanced citywide tuberculosis surveillance and spatial analysis"

subtitle: "Spatial analysis"

author: |
  | Peter MacPherson
  |
  | Liverpool School of Tropical Medicine, Liverpool, UK
  | Malawi-Liverpool-Wellcome Clinical Research Programme, Blantyre, Malawi
  |

date: | 
  | `r format(Sys.time(), "%B %d, %Y")`
  |
  | Table of Contents:
output: 
  html_document:
    theme: cosmo
    highlight: espresso
    toc: true
---

##1. Set-up

Load all the required libraries.

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(sf)
library(ggthemes)
library(viridis)
library(scales)
library(lubridate)
library(brms)
library(arsenal)
library(knitr)
library(here)
library(spdep)
library(broom)
library(RANN)
library(forcats)
library(cowplot)
library(bayesplot)
```

<br>

##2. Reproducibility

This analysis was run on *`r Sys.time()`*, using the following system profile and package versions:

```{r sysinfo, echo=TRUE, results='markup', message=FALSE, comment=NA, warning=FALSE}
sessionInfo()
```

<br>

##3. Map theme

Create a base theme function for the maps.

```{r,warning=FALSE}
map_light <- function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "Helvetica", color = "#22211d"),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(color = "#ebebe5", size = 0.2),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA), 
    panel.background = element_rect(fill = "white", color = NA), 
    legend.background = element_rect(fill = "white", color = NA),
    panel.border = element_blank(),
    ...
  )
}

```

<br>

##4. Total population

Describe summary characteristics of the population.

```{r,warning=FALSE}
data("dat", package="BlantyreTBEpi")

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(total=sum(total))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(min=min(total))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(min=max(total))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(mean=mean(total),
            sd= sd(total))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(median=median(total),
            i25=quantile(total, probs = 0.25),
            i75=quantile(total, probs = 0.75))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(min_m_f_adult_ratio = min(m_f_adult_ratio),
            mean_m_f_adult_ratio = mean(m_f_adult_ratio),
            max_m_f_adult_ratio = max(m_f_adult_ratio))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(min_prop_adults = min(prop_adults),
            mean_prop_adults = mean(prop_adults),
            max_prop_adults = max(prop_adults))

dat %>%
  st_set_geometry(., NULL) %>%
  summarise(min_pov = min(mwi11povcons200),
            mean_pov = mean(mwi11povcons200),
            max_pov = max(mwi11povcons200))


```


Plot the total population of Blantyre.

```{r,warning=FALSE}
data("dat", package="BlantyreTBEpi")
data("clinics", package="BlantyreTBEpi")

g1<- ggplot() +
  geom_sf(data = dat, aes(fill = total)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name ="Count",
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g1

ggsave(file="Figure1.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##5. Population density

Plot the population density of Blantyre.


```{r,warning=FALSE}
g2 <- ggplot() +
  geom_sf(data = dat, aes(fill = popdens)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Population\nper\nsquare km"),
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g2

ggsave(file="popdensity.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##6. People per household

Plot the mean people per household Blantyre.


```{r,warning=FALSE}
g3 <- ggplot() +
  geom_sf(data = dat, aes(fill = mean_people_per_hh)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Mean number\nof people\nper household"),
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g3

ggsave(file="mean_hh_people.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##7. Poverty distribution

Plot the estimaed mean proportion of the population living on less than $2 per day. [Note this is estimated from Worldpop data]


```{r,warning=FALSE}
  
g4 <- ggplot() +
  geom_sf(data = dat, aes(fill = mwi11povcons200)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Mean percentage\nof people\nliving on\nless than\nUS $2 per day"),
                     labels=comma,
                     limits = c(0,100)) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g4

ggsave(file="poverty.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##8. Distance to nearest TB registration site

```{r,warning=FALSE}
data("dat", package="BlantyreTBEpi")

g5 <- ggplot() +
  geom_sf(data = dat, aes(fill = clinic_distance)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Distance (m)\nto nearest\nTB registration\ncentre"),
                     labels=comma,
                     limits = c(0,10000)) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g5

ggsave(file="distance_to_clinic.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```


<br>

##9. Male to female adult ratio


```{r,warning=FALSE}
g6 <- ggplot() +
  geom_sf(data = dat, aes(fill = m_f_adult_ratio)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Adult\nMale:Female\nratio"),
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g6

ggsave(file="m_f_ratio.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##10. Percentage of HSA cluster population who is adult


```{r,warning=FALSE}
g7 <- ggplot() +
  geom_sf(data = dat, aes(fill = prop_adults)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Percent\naged\n15 years\nor older"),
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g7

ggsave(file="prop_adults.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))

```

<br>

##11. Sputum smear positive to negative ratio.


```{r}
#Add in a term for ratio of smearpositive to smearnegative cases

dat <- dat %>%
  mutate(smr_ratio = n_smearpos_cases/(n_cases-n_smearpos_cases))

dat$smr_ratio[is.nan(dat$smr_ratio)] <- 1
dat$smr_ratio[is.infinite(dat$smr_ratio)] <- 1


g8 <- ggplot() +
  geom_sf(data = dat, aes(fill = smr_ratio)) +
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_viridis(option = "magma", direction = -1,
                     name = expression ("Smr +ve:-ve\nratio"),
                     labels=comma) +
  labs(x = NULL, 
       y = NULL) +
  theme(legend.text.align = 1)

g8

ggsave(file="smr_pos_neg.pdf", width=6, height=6, dpi=300, 
       path = here("figures"))




```

<br>

##12. Supplemental Figure 1

Join all the descriptive figures together to produce one figure

```{r,warning=FALSE}
plot_grid(g1, g2, g3, g4, g5, g6, g7, g8,
          labels = c("A", "B", "C", "D", "E", "F", "G", "H"),
          label_fontfamily = "Helvetica")

ggsave(file="S1_Fig.pdf", width=12, height=12, dpi=300, 
       path = here("figures"))

```


##13. Correlation between variables.

```{r}

corrs <- dat %>%
  st_set_geometry(., NULL) %>%
  dplyr::select(total, mean_people_per_hh, popdens, m_f_adult_ratio,
         prop_adults, mwi11povcons200, clinic_distance, smr_ratio)
   

pairs(corrs)

```


<br>

##14. Table 1: Characteristics of TB cases

Compare characteristics of cases located within and outside of HSA catchment areas.

```{r,warning=FALSE, results="asis"}

data("all_cases_sf", package="BlantyreTBEpi")

all_cases_sf$year <- fct_drop(all_cases_sf$year)


table1 <- tableby(resident ~ year + sex + age + tbclass + smearstatus + hivstatus +
                    lab_smear + lab_culture + lab_id + any_micro_confirm, data = all_cases_sf)

summary(table1)
```


<br>
 
##16. Figure 2: TB case notification rates
 
Plot overall TB case notification rates.

```{r,warning=FALSE}

data("cnrs", package="BlantyreTBEpi")

cnr1 <- ggplot() +
  geom_sf(data=dat) +
  geom_sf(data=cnrs, aes(fill=cnr)) + 
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_distiller(palette = "Spectral",
                     name = "",
                     labels=comma,
                     limits = c(0,999)) +
  labs(x = NULL, 
       y = NULL) +
  facet_grid(~year, drop = FALSE) +
  theme(panel.background = element_rect(fill = "white", color = "#22211d" )) 

cnr1

ggsave(file="cnrs.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))

```

Now plot microbiologically-confirmed TB case notification rate

```{r,warning=FALSE}

data("cnrs", package="BlantyreTBEpi")

cnr2 <- ggplot() +
  geom_sf(data=dat) +
  geom_sf(data=cnrs, aes(fill=micro_cnr)) + 
  geom_sf(data=clinics, shape=17, colour="#22211d") +
  map_light() +
  scale_fill_distiller(palette = "Spectral",
                     name = "",
                     labels=comma,
                     limits = c(0,500)) +
  labs(x = NULL, 
       y = NULL) +
  facet_grid(~year, drop = FALSE) +
  theme(panel.background = element_rect(fill = "white", color = "#22211d" ))

cnr2

ggsave(file="micro_cnrs.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))

```

```{r,warning=FALSE}
plot_grid(cnr1, cnr2, labels = c("A", "B"), ncol = 1,
          label_fontfamily = "Helvetica")

ggsave(file="Figure2.pdf", width=10, height=8, dpi=300, 
       path = here("figures"))


```



##17.Spatial modelling

Make the spatial correlation matrices.

```{r, warning=FALSE}
#Make the adjacency matrix
#First find the centroids
centroids_sf <- st_centroid(st_geometry(dat), of_largest_polygon=TRUE)
zm <- class(st_geometry(centroids_sf)[[1]])[1]
if (zm %in% c("XYM", "XYZM"))
  centroids_sf <- st_zm(centroids_sf, drop=TRUE, what="ZM")
if (zm %in% c("XYZ"))
  centroids_sf <- st_zm(centroids_sf, drop=TRUE, what="ZM")

coords_sf <- st_coordinates(centroids_sf)

neigh1 <- knn2nb(knearneigh(coords_sf, k=1))
neigh2 <- knn2nb(knearneigh(coords_sf, k=2))
neigh3 <- knn2nb(knearneigh(coords_sf, k=3))
neigh4 <- knn2nb(knearneigh(coords_sf, k=4))
neigh5 <- knn2nb(knearneigh(coords_sf, k=5))
neigh6 <- knn2nb(knearneigh(coords_sf, k=6))


path <- paste(here("figures"), "S2_fig.pdf", sep="/")
pdf(path)

par(mfrow=c(2,3), mar=c(1,1,1,1))
plot.nb(neigh1, coords_sf, col="#036564", points=FALSE)
title(main="k=1", cex=2)
plot.nb(neigh2, coords_sf, col="#036564", points=FALSE)
title(main="k=2", cex=2)
plot.nb(neigh3, coords_sf, col="#036564", points=FALSE)
title(main="k=3", cex=2)
plot.nb(neigh4, coords_sf, col="#036564", points=FALSE)
title(main="k=4", cex=2)
plot.nb(neigh5, coords_sf, col="#036564", points=FALSE)
title(main="k=5", cex=2)
plot.nb(neigh6, coords_sf, col="#036564", points=FALSE)
title(main="k=6", cex=2)

dev.off()

dat$hsa_area <- factor (1:nrow (dat))

#Create the neighbourhood matrices for each model
w1 <- nb2mat(make.sym.nb(neigh1), style="B")
rownames(w1) <- as.character (1:nrow (dat))

w2 <- nb2mat(make.sym.nb(neigh2), style="B")
rownames(w2) <- as.character (1:nrow (dat))

w3 <- nb2mat(make.sym.nb(neigh3), style="B")
rownames(w3) <- as.character (1:nrow (dat))

w4 <- nb2mat(make.sym.nb(neigh4), style="B")
rownames(w4) <- as.character (1:nrow (dat))

w5 <- nb2mat(make.sym.nb(neigh5), style="B")
rownames(w5) <- as.character (1:nrow (dat))

w6 <- nb2mat(make.sym.nb(neigh6), style="B")
rownames(w6) <- as.character (1:nrow (dat))



```

Take logarithm of some values

```{r}

dat <- dat %>%
  mutate(log_total = log(total))

dat <- dat %>%
  mutate(log_popdens = log(popdens))

dat <- dat %>%
  mutate(log_clinic_distance = log(clinic_distance))


```


Run all of the models, varying spatial autocorration prior


```{r}

prior <- c(prior_string("normal(0,10)", class="b"),
           prior_(~normal(0,10), class= ~Intercept))

all_0 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_1 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w2, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_2 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w2, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))


all_3 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w3, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_4 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w4, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_5 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w5, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_6 <- brm(bf(n_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w6, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

all_0_ic <- waic(all_0)
all_1_ic <- waic(all_1)
all_2_ic <- waic(all_2)
all_3_ic <- waic(all_3)
all_4_ic <- waic(all_4)
all_5_ic <- waic(all_5)
all_6_ic <- waic(all_6)

compare_ic(all_0_ic, all_1_ic, all_2_ic, all_3_ic, all_4_ic, all_5_ic, all_6_ic)

waics0 <- as.tibble(unlist(all_0_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k0") %>%
  dplyr::select(model, value)

waics1 <- as.tibble(unlist(all_1_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k1") %>%
  dplyr::select(model, value)

waics2 <- as.tibble(unlist(all_2_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k2") %>%
  dplyr::select(model, value)

waics3 <- as.tibble(unlist(all_3_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k3") %>%
  dplyr::select(model, value)

waics4 <- as.tibble(unlist(all_4_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k4") %>%
  dplyr::select(model, value)

waics5 <- as.tibble(unlist(all_5_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k5") %>%
  dplyr::select(model, value)

waics6 <- as.tibble(unlist(all_6_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k6") %>%
  dplyr::select(model, value)

waics <- bind_rows(waics0, waics1, waics2, waics3, waics4, waics5, waics6)
waics <- waics %>% mutate(value = as.numeric(value)) %>% mutate(value = round(value, digits=2))

ests_all_0<- tidy(exp(fixef(all_0)))
ests_all_0 <- ests_all_0 %>%
  mutate(model="k0")

ests_all_1<- tidy(exp(fixef(all_1)))
ests_all_1 <- ests_all_1 %>%
  mutate(model="k1")

ests_all_2<- tidy(exp(fixef(all_2)))
ests_all_2 <- ests_all_2 %>%
  mutate(model="k2")

ests_all_3<- tidy(exp(fixef(all_3)))
ests_all_3 <- ests_all_3 %>%
  mutate(model="k3")

ests_all_4<- tidy(exp(fixef(all_4)))
ests_all_4 <- ests_all_4 %>%
  mutate(model="k4")

ests_all_5<- tidy(exp(fixef(all_5)))
ests_all_5 <- ests_all_5 %>%
  mutate(model="k5")

ests_all_6<- tidy(exp(fixef(all_6)))
ests_all_6 <- ests_all_6 %>%
  mutate(model="k6")

all_all <- bind_rows(ests_all_0, ests_all_1, ests_all_2, ests_all_3, ests_all_4, ests_all_5, ests_all_6)

all_all <- left_join(all_all, waics)

all_all <- all_all %>% mutate(value = as.numeric(value)) %>% mutate(value = round(value, digits=2))

s3a <- all_all %>%
  filter(`.rownames` != "Intercept") %>%
  ggplot() +
  geom_pointrange(aes(x=.rownames, y=Estimate, ymin=Q2.5, ymax=Q97.5, group=fct_rev(model), colour=model),
                  position=position_dodge(width=1)) +
  coord_flip() +
  geom_hline(yintercept = 1, linetype="dashed") +
  ylab("") +
  xlab("") +
  scale_x_discrete(labels = c("Log(10) distance to clinic",
                              "Log(10) population density",
                              "M:F adult ratio",
                              "Mean people per household",
                              "Percentage living in poverty",
                              "Percentage adults",
                              "Smear pos:smear neg ratio")) +
  scale_color_economist(name = "Model: WAIC", labels = paste(levels(as.factor(all_all$model)), waics$value, sep=": ")) +
  scale_y_sqrt(breaks=c(0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6), limits=c(0,1.6)) +
  theme_bw() +
  theme(text =  element_text(size=14))

ggsave(file="s3a_fig.pdf", height = 10, width = 10, dpi=300, 
       path = here("figures"))


```



```{r}

prior <- c(prior_string("normal(0,10)", class="b"),
           prior_(~normal(0,10), class= ~Intercept))

micro_0 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_1 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w2, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_2 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w2, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))


micro_3 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w3, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_4 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w4, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_5 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w5, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_6 <- brm(bf(n_micro_cases ~
          mean_people_per_hh +
          log_popdens +
          log_clinic_distance +
          mwi11povcons200 +
          m_f_adult_ratio +
          prop_adults +
          smr_ratio +
          offset(log_total)),
          data=dat, 
          family='poisson',
          prior = prior,
          autocor=cor_car(w6, ~ 1 | c02hsaid1),
          iter=4000, warmup=1000,
          chains=3,
          seed=1234,
          control=list(adapt_delta=0.95))

micro_0_ic <- waic(micro_0)
micro_1_ic <- waic(micro_1)
micro_2_ic <- waic(micro_2)
micro_3_ic <- waic(micro_3)
micro_4_ic <- waic(micro_4)
micro_5_ic <- waic(micro_5)
micro_6_ic <- waic(micro_6)

compare_ic(micro_0_ic, micro_1_ic, micro_2_ic, micro_3_ic, micro_4_ic, micro_5_ic, micro_6_ic)

waics_micro_0 <- as.tibble(unlist(micro_0_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k0") %>%
  dplyr::select(model, value)

waics_micro_1 <- as.tibble(unlist(micro_1_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k1") %>%
  dplyr::select(model, value)

waics_micro_2 <- as.tibble(unlist(micro_2_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k2") %>%
  dplyr::select(model, value)

waics_micro_3 <- as.tibble(unlist(micro_3_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k3") %>%
  dplyr::select(model, value)

waics_micro_4 <- as.tibble(unlist(micro_4_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k4") %>%
  dplyr::select(model, value)

waics_micro_5 <- as.tibble(unlist(micro_5_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k5") %>%
  dplyr::select(model, value)

waics_micro_6 <- as.tibble(unlist(micro_6_ic)) %>% rownames_to_column() %>% filter(rowname == "waic") %>%
  mutate(model="k6") %>%
  dplyr::select(model, value)

waics_micro <- bind_rows(waics_micro_0, waics_micro_1, waics_micro_2, waics_micro_3, waics_micro_4, waics_micro_5, waics_micro_6)

waics_micro <- waics_micro %>% mutate(value = as.numeric(value)) %>% mutate(value = round(value, digits=2))


ests_micro_0<- tidy(exp(fixef(micro_0)))
ests_micro_0 <- ests_micro_0 %>%
  mutate(model="k0")

ests_micro_1<- tidy(exp(fixef(micro_1)))
ests_micro_1 <- ests_micro_1 %>%
  mutate(model="k1")

ests_micro_2<- tidy(exp(fixef(micro_2)))
ests_micro_2 <- ests_micro_2 %>%
  mutate(model="k2")

ests_micro_3<- tidy(exp(fixef(micro_3)))
ests_micro_3 <- ests_micro_3 %>%
  mutate(model="k3")

ests_micro_4<- tidy(exp(fixef(micro_4)))
ests_micro_4 <- ests_micro_4 %>%
  mutate(model="k4")

ests_micro_5<- tidy(exp(fixef(micro_5)))
ests_micro_5 <- ests_micro_5 %>%
  mutate(model="k5")

ests_micro_6<- tidy(exp(fixef(micro_6)))
ests_micro_6 <- ests_micro_6 %>%
  mutate(model="k6")

micro_micro <- bind_rows(ests_micro_0, ests_micro_1, ests_micro_2, ests_micro_3, ests_micro_4, ests_micro_5, ests_micro_6)

micro_micro <- left_join(micro_micro, waics_micro)

micro_micro <- micro_micro %>% mutate(value = as.numeric(value))

s3b <- micro_micro %>%
  filter(`.rownames` != "Intercept") %>%
  ggplot() +
  geom_pointrange(aes(x=.rownames, y=Estimate, ymin=Q2.5, ymax=Q97.5, group=fct_rev(model), colour=model),
                  position=position_dodge(width=1)) +
  coord_flip() +
  geom_hline(yintercept = 1, linetype="dashed") +
  ylab("") +
  xlab("") +
  scale_x_discrete(labels = c("Log(10) distance to clinic",
                              "Log(10) population density",
                              "M:F adult ratio",
                              "Mean people per household",
                              "Percentage living in poverty",
                              "Percentage adults",
                              "Smear pos:smear neg ratio")) +
  scale_color_economist(name = "Model: WAIC", labels = paste(levels(as.factor(all_all$mode)), sprintf("%.2f", waics_micro$value), sep=": ")) +
  scale_y_sqrt(breaks=c(0.2,0.4,0.6,0.8,1.0,1.2,1.4,1.6), limits=c(0,1.6)) +
  theme_bw() +
  theme(text =  element_text(size=14))

ggsave(file="s3b_fig.pdf", width = 10, height = 10, dpi=300, 
       path = here("figures"))


```


```{r}

plot_grid(s3a, s3b,
          labels = c("A", "B"), ncol = 1,
          label_fontfamily = "Helvetica")

ggsave(file="S3_fig.pdf", width=10, height=12, dpi=300, 
       path = here("figures"))


``` 






All TB cases final model

Select model with lowest WAIC (i.e. k6)

```{r}

#Overall model summary
summary(all_6)

#Model diagnostics
plot(all_6, ask = FALSE)

pp_all_6 <- as_tibble(predict(all_6))
pp_all_6 <- dat %>%
  dplyr::select(c02hsaid, n_cases) %>%
  bind_cols(pp_all_6) %>%
  mutate(diff=n_cases - Estimate) %>%
  mutate(abdiff = abs(diff)) %>%
  arrange(diff) %>%
  mutate(c02hsaid = factor(c02hsaid))

residual1 <- pp_all_6 %>%
  ggplot() +
  geom_linerange(aes(x=fct_reorder(c02hsaid, n_cases, .desc=TRUE), ymin=n_cases, ymax=Estimate), alpha=0.7) +
  geom_point(aes(x=fct_reorder(c02hsaid, n_cases, .desc=TRUE), y=n_cases), colour="firebrick") +
  geom_point(aes(x=fct_reorder(c02hsaid, n_cases, .desc=TRUE), alpha=abdiff, y=Estimate), colour="steelblue") +
  theme_classic() +
  labs(x="",
       y="Observed (red) and predicted (blue) TB cases") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  theme(legend.position="none")

ggsave(file="residual1.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))


##Model estimates
ests_all_6<- tidy(exp(fixef(all_6)))

ests_all_6 %>%
  kable()


#Plot posterior distributions
facet_names <- c(
                    `b_log_popdens` = "Population density (per square km)",
                    `b_mean_people_per_hh` = "Mean people per household",
                    `b_log_clinic_distance` = "Distance to TB clinic",
                    `b_mwi11povcons200` = "Percentage living in poverty",
                    `b_m_f_adult_ratio` = "Male:female adult ratio",
                    `b_prop_adults` = "Percentage adults",
                    `b_smr_ratio` = "Smear pos:smear neg ratio")

plotpars <- c("b_log_popdens", "b_mean_people_per_hh", "b_log_clinic_distance",
              "b_mwi11povcons200", "b_m_f_adult_ratio", "b_prop_adults", 
              "Smear pos:smear neg ratio")

post_samples_all_6 <- posterior_samples(all_6, "b")
plot_post_all_6 <- gather(post_samples_all_6, variable, value) %>%
  filter(variable %in% plotpars)
plot_post_all_6$variable <- factor(plot_post_all_6$variable)

plot_post_all_6$variable <- fct_relevel(plot_post_all_6$variable,"b_Intercept")

plot_post_all_6 %>% 
  filter(variable %in% plotpars) %>%
  ggplot(aes(value)) +
  geom_density(fill="firebrick", colour="firebrick", alpha=0.5) +
  facet_wrap(~variable, nrow = 2, labeller = as_labeller(facet_names)) +
  geom_vline(xintercept = 0, linetype="dashed") +
  theme_bw() +
  xlim(-5,5) +
  xlab("Estimated effect on TB case notification rate") +
  ylab("") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) -> post_all_6

post_all_6


ggsave(file="posterior1.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))


```



Select model with lowest WAIC for microbiologically-confirmed TB (i.e. k4).

```{r}

#Overall model summary
summary(micro_4)

#Model diagnostics
plot(micro_4, ask = FALSE)

Resid.zip.brm <- residuals(micro_4, type='pearson')[,'Estimate']
Fitted.zip.brm <- exp(fitted(micro_4, scale='linear')[1:nrow(dat),'Estimate'])
ggplot(data=NULL, aes(y=Resid.zip.brm, x=Fitted.zip.brm)) + geom_point()

pp_micro_4 <- as_tibble(predict(micro_4))
pp_micro_4 <- dat %>%
  dplyr::select(c02hsaid, n_micro_cases) %>%
  bind_cols(pp_micro_4) %>%
  mutate(diff=n_micro_cases - Estimate) %>%
  mutate(abdiff = abs(diff)) %>%
  arrange(diff) %>%
  mutate(c02hsaid = factor(c02hsaid))

residual2 <- pp_micro_4 %>%
  ggplot() +
  geom_point(aes(x=fct_reorder(c02hsaid, n_micro_cases, .desc=TRUE), y=n_micro_cases), colour="firebrick") +
  geom_point(aes(x=fct_reorder(c02hsaid, n_micro_cases, .desc=TRUE), alpha=abdiff, y=Estimate), colour="steelblue") +
  geom_linerange(aes(x=fct_reorder(c02hsaid, n_micro_cases, .desc=TRUE), ymin=n_micro_cases, ymax=Estimate), alpha=0.7) +
  theme_classic() +
  labs(x="",
       y="Observed (red) and predicted (blue) TB cases") +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  theme(legend.position="none")

ggsave(file="residual2.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))


##Model estimates
ests_micro_4<- tidy(exp(fixef(micro_4)))

ests_micro_4 %>%
  kable()


#Plot posterior distributions
facet_names <- c(
                    `b_log_popdens` = "Population density (per square km)",
                    `b_mean_people_per_hh` = "Mean people per household",
                    `b_log_clinic_distance` = "Distance to TB clinic",
                    `b_mwi11povcons200` = "Percentage living in poverty",
                    `b_m_f_adult_ratio` = "Male:female adult ratio",
                    `b_prop_adults` = "Percentage adults",
                    `b_smr_ratio` = "Smear pos:smear neg ratio")

plotpars <- c("b_log_popdens", "b_mean_people_per_hh", "b_log_clinic_distance",
              "b_mwi11povcons200", "b_m_f_adult_ratio", "b_prop_adults", "Smear pos:smear neg ratio")

post_samples_micro_4 <- posterior_samples(micro_4, "b")
plot_post_micro_4 <- gather(post_samples_micro_4, variable, value) %>%
  filter(variable %in% plotpars)
plot_post_micro_4$variable <- factor(plot_post_micro_4$variable)

plot_post_micro_4$variable <- fct_relevel(plot_post_micro_4$variable,"b_Intercept")

plot_post_micro_4 %>% 
  filter(variable %in% plotpars) %>%
  ggplot(aes(value)) +
  geom_density(fill="firebrick", colour="firebrick", alpha=0.5) +
  facet_wrap(~variable, nrow = 2, labeller = as_labeller(facet_names)) +
  geom_vline(xintercept = 0, linetype="dashed") +
  theme_bw() +
  xlim(-5,5) +
  xlab("Estimated effect on TB case notification rate per unit increase in variable") +
  ylab("") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()) -> post_micro_4

post_micro_4


ggsave(file="posterior2.pdf", width=12, height=6, dpi=300, 
       path = here("figures"))


```

<br>

## 18. Figure 3: Marginal Effects plot

```{r}
me <- marginal_effects(all_6, 
                       conditions = data.frame(total = seq(1000,15000, 250)))
me2 <- marginal_effects(micro_4, conditions = data.frame(total = seq(1000,15000, 250)))

p1 <- me[[1]]
p1 <- p1 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=mean_people_per_hh) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Mean people per household") 

p2 <- me[[2]]
p2 <- p2 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=log_popdens) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Log population density")

p3 <- me[[3]]
p3 <- p3 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=log_clinic_distance) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Log clinic distance")

p4 <- me[[4]]
p4 <- p4 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=mwi11povcons200) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Percentage living in poverty")

p5 <- me[[5]]
p5 <- p5 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=m_f_adult_ratio) %>% 
  mutate(type="All TB") %>%
  mutate(param = "M:F adult ratio")

p6 <- me[[6]]
p6 <- p6 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=prop_adults) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Percentage adults")

p7 <- me[[7]]
p7 <- p7 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=smr_ratio) %>% 
  mutate(type="All TB") %>%
  mutate(param = "Smear positive:negative ratio")

p8 <- me2[[1]]
p8 <- p8 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=mean_people_per_hh) %>% 
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Mean people per household")

p9 <- me2[[2]]
p9 <- p9 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=log_popdens) %>% 
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Log population density")

p10 <- me2[[3]]
p10 <- p10 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=log_clinic_distance) %>% 
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Log clinic distance")

p11 <- me2[[4]]
p11 <- p11 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=mwi11povcons200) %>% 
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Percentage living in poverty")

p12 <- me2[[5]]
p12 <- p12 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=m_f_adult_ratio) %>%
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "M:F adult ratio")

p13 <- me2[[6]]
p13 <- p13 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=prop_adults) %>%
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Percentage adults")

p14 <- me2[[7]]
p14 <- p14 %>%
  mutate(rate=(estimate__/total)*100000) %>%
  mutate(l95 = (lower__/total)*100000) %>%
  mutate(u95 = (upper__/total)*100000) %>%
  select(rate, l95,u95, value=smr_ratio) %>%
  mutate(type="Microbiologically-confirmed TB") %>%
  mutate(param = "Smear positive:negative ratio")

pz <- bind_rows(p1, p2, p3, p4, p5, p6, p7,
                p8, p9, p10, p11, p12, p13, p14)


ggplot(pz) +
  geom_line(aes(x=value, y=rate, group=type, colour=type)) +
  geom_ribbon(aes(x=value, ymin=l95, ymax=u95, group=type, fill=type), alpha=0.3) +
  facet_wrap(~param, scales = "free") +
  ylim(0,1200) +
  ylab("TB case notification rate per 100,000 (95% credible interval)") +
  xlab("") +
  theme_minimal() +
  theme(legend.position="none")

ggsave(file="Figure3.pdf", 
       path = here("figures"))

```




##19. Supplemental Figure 3: Predicted and observed TB case notifications

```{r}
plot_grid(residual1, residual2,
          labels = c("A", "B"), ncol = 1,
          label_fontfamily = "Helvetica")

ggsave(file="S3_Fig.pdf", width=8, height=8, dpi=300, 
       path = here("figures"))

```


